The CrypTool Script: \\ Cryptography, Mathematics and More \\ \vspace{30pt} {\large Background reading\\%*[-8pt] for the free e-learning program CrypTool\\ (10th edition -- distributed with CrypTool version 1.4.30)} \vspace{60pt}

### 2012-10-11 script/en/chapters/numbertheory.tex
Small corrections and enhancements in the CT Script

% $Id$
% \def\QM {{,\kern -0.9 pt ,}}
\setcounter{theorem}{0}
\setcounter{definition}{0}

% ..........................................................................
% --------------------------------------------------------------------------
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%              E l e m e n t a r e  Z a h l e n t h e o r i e
% /~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\newpage
\hypertarget{Chapter_ElementaryNT}{}
\chapter{Introduction to Elementary Number Theory with Examples}
\label{Chapter_ElementaryNT}
(Bernhard Esslinger, July 2001; Updates: Dec. 2001, June 2002, May 2003,
 May 2005, March 2006, June 2007, July 2009, Jan. 2010) \\

@@

@@

\begin{center}
\begin{tabular}{|l|l|l|}\hline
 & Short description ~~ & Page \\ \hline
Definition~\ref{def-zth-prime} & prime numbers &  \pageref{def-zth-prime} \\
Definition~\ref{def-zth-composite} & composite numbers & \pageref{def-zth-composite}  \\ \hline
Theorem~\ref{thm-zth-cnum} & factors of composite numbers~~~~~~~ & \pageref{thm-zth-cnum}\\
Theorem~\ref{thm-zth-mthm} &  1st fundamental theorem of number theory &  \pageref{thm-zth-mthm} \\  \hline
Definition~\ref{def-zth-divisibility} & divisibility & \pageref{def-zth-divisibility} \\
Definition~\ref{def-zth-remainder} & remainder class $r$ modulo $m$ & \pageref{def-zth-remainder} \\
Definition~\ref{def-zth-congruence} & congruent & \pageref{def-zth-congruence} \\ \hline
Theorem~\ref{thm-zth-div} & congruence with difference  & \pageref{thm-zth-div} \\
Theorem~\ref{thm-zth-multinv} & multiplicative inverse (existence) & \pageref{thm-zth-multinv}  \\
Theorem~\ref{thm-zth-exhperm} & exhaustive permutation & \pageref{thm-zth-exhperm} \\
Theorem~\ref{thm-zth-pot} & power mod $m$ & \pageref{thm-zth-pot} \\ \hline
Definition~\ref{def-zth-zn} & $\mathbb{Z}_n$  & \pageref{def-zth-zn}\\
Definition~\ref{def-zth-znmult} &   $\mathbb{Z}_n^*$ & \pageref{def-zth-znmult} \\ \hline
Theorem~\ref{thm-zth-znmult} & multiplicative inverse in $\mathbb{Z}_n^*$& \pageref{thm-zth-znmult} \\ \hline
Definition~\ref{def-zth-phiofn} & Euler function $J(n)$ & \pageref{def-zth-phiofn} \\
Theorem~\ref{thm-zth-phiprime} & $J(p)$ &  \pageref{thm-zth-phiprime}\\
Theorem~\ref{thm-zth-phipq} & $J(p*q)$ &  \pageref{thm-zth-phipq}\\
Theorem~\ref{thm-zth-phimultprime} & $J(p_1 * \cdots *p_k)$ & \pageref{thm-zth-phimultprime} \\
Theorem~\ref{thm-zth-phinum} & $J(p_1^{e_1} * \cdots *p_k^{e_k})$ & \pageref{thm-zth-phinum} \\
Theorem~\ref{thm-zth-fermat1} & little Fermat  &  \pageref{thm-zth-fermat1}\\
Theorem~\ref{thm-zth-fermateuler} & Euler-Fermat theorem & \pageref{thm-zth-fermateuler} \\ \hline
Definition~\ref{def-zth-ordn} & multiplicative order $ {\rm ord}_{m} (a)$ & \pageref{def-zth-ordn} \\
Definition~\ref{def-zth-primitiveroot} & primitive root of $m$ &  \pageref{def-zth-primitiveroot}\\
Theorem~\ref{thm-zth-ordp} & exhausting of all possible values & \pageref{thm-zth-ordp} \\ \hline
\end{tabular}
\end{center}
\vskip +6 pt

@@

\bibitem[Agrawal2002]{nt:Agrawal2002}  \index{Agrawal 2002} 
    M. Agrawal, N. Kayal, N. Saxena, \\
    {\em PRIMES in P}, August 2002, revised paper: \\
       \url{http://www.cse.iitk.ac.in/~manindra/algebra/primality_v6.pdf}\\
    See also the website "The AKS "PRIMES in P" Algorithm Resource":
       \url{http://fatphil.org/maths/AKS/}.

@@

\bibitem[Bourseau2002]{nt:Bourseau2002} \index{Bourseau 2002} \index{Fox 2002}
    F. Bourseau, D. Fox, C. Thiel, \\
    {\em Vorz\"uge und Grenzen des RSA-Verfahrens},\\
    In: Datenschutz und Datensicherheit (DuD) 26/2002, pp~84-89 (see www.dud.de),
    \url{http://www.secorvo.de/publikationen/rsa-grenzen-fox-2002.pdf}.

@@

  \item {Henrik Koy for making many very useful suggestions, 
         for the very constructive proof-reading of the first version of
	 this article, and for helping with TeX. \\
	 Mr.\ Koy designed and developed in his leisure time the 
	 functions and the complex dialog box of the RSA cryptosystem within CrypTool, 
	 which enables you to execute the RSA samples of this article.}

@@

  \item {Minh Van Nguyen from Australia for his always fast, professional and exhaustive
         help with the Sage\index{Sage} code samples in this chapter.}

>>>>

% $Id$
% ..........................................................................
% --------------------------------------------------------------------------
% ++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++++
%              E l e m e n t a r e  Z a h l e n t h e o r i e
% ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

% \def\QM {{,\kern -0.9 pt ,}}
\setcounter{theorem}{0}
\setcounter{definition}{0}

\newpage
\hypertarget{Chapter_ElementaryNT}{}
\chapter{Introduction to Elementary Number Theory with Examples}
\label{Chapter_ElementaryNT}
(Bernhard Esslinger, July 2001; Updates: Dec. 2001, June 2002, May 2003,
 May 2005, March 2006, June 2007, July 2009, Jan. 2010, Sep. 2012) \\

@@

% ---------------------------------------------------------------------------
\clearpage
\newpage
\hypertarget{NumberTheory_Sage_Number-of-RSA-FixedPoints}{}
\subsection
    [RSA fixed points \texorpdfstring{}{m = m\^{}e}]
    {RSA fixed points $ m^e = m \bmod n $ mit $m \in \{1,...,n-1\}$ }
\label{l:NumberTheory_Sage_Number-of-RSA-FixedPoints}{}
\index{Fixpoint}\index{RSA!fixpoint}
%%% xxx111222xxx-beg

Also encryption methods can have fixed -- cleartext messages where the
according ciphertext matches the original. In mathematics, variables
mapped by the algorithm (function) onto themselves are called fixed points.
In cryptography the according messages are called ``unconcealed messages''.

Generally speaking: The more fixed points an encryption algorithm contains,
the easier it is to break it.

With the RSA procedure: $n=pq$ is the product of two different prime numbers,
and there exists $e$ where $gcd(e,(p-1)(q-1))=1$. The encription is then
$c = m^e \mod n$. 
A fixed point in the RSA procedure is the message $m$, where: 
$m = m^e \mod n$. 
The result of the encryption is the given message.

When the size of $n$ is sufficiently big, the probability of the occurance of fixed
points in RSA is very small -- as illustrated in Figure \ref{fig:NumberFixpointsGrowingN}:
In average, we found not more than 40 fixed points.

Students often presume the occurence of fixed points high, because they counter
a ``relatively'' large number of examples when experimenting with \textbf{small}
prime numbers, as m = 0, 1 and n-1 are also always fixed points.

In practice, where large prime numbers are chosen, fixed points have no
significance for the security of RSA. Therefore, this paragraph refers more
to the mathematical questions\footnote{%
Thanks to Taras Shevchenko for gathering parts of the content of this chapter
and to Volker Simon for writing the Sage program
\ref{nt_sagesample_Calculate_RSA-Fixpoints} ``Getfixpoints''.}.

% ----------------------------------------------------
\subsubsection{The number of RSA fixed points}

In this section we show how many RSA fixed points there are for $m \in \{1,...,n-1\} $.\\

\begin{theorem}\label{nt-number-of-fixpoints-1-to-n-1}
The number of the fixed points $ m^e = m \bmod n$ with $m \in \{1,...,n-1\} $ is \\$ gcd(p-1, e-1) \cdot gcd(q-1, e-1) $.
\end{theorem}
\begin{proof}
Given $m^e = m \bmod n.$
According to the CRT, the following statements are equivalent:\\
$$ [ m^e = m \bmod n ]  \Leftrightarrow [ m^{e} = m \bmod p \text{  and  } m^{e} = m \bmod q $$
% STH: consider using 'analysis' instead of 'decomposition'
%      be: Klären, denn es ist auch keine Äquivalenz! xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
This decomposition is equivalent to:
$$m^{e-1} = 1 \bmod p \text{  and  } m^{e-1} = 1 \bmod q. $$

\noindent We consider  $m^{e-1} = 1 \bmod p$ and search all $(e-1)$
roots of unity\index{RootOfUnity}\footnote{%
- In mathematics, a \textbf{root of unity} is a number $x$ that equals 1 when raised to some integer power $n$.

\noindent- An $n$-th root of unity $x$ is \textbf{primitive} if it is not a $k$-th root of unity for all integers $k$ smaller than $n$:
 $$x^{n} = 1  \text{ and }    x^{k} \neq 1 ~~~(k = 1,2, 3, ..., n-1)$$

\noindent-
If $F$ is a finite field and $n$ is a positive integer, then a $n$th-root of unity in $F$ is a solution of the equation $$ x^{n}-1 = 0 \text{ in } F $$
}
in $\mathbb{Z}_p^{*}.$\\
It holds: $\mathbb{Z}_p^{*}$ for p prime is cyclic.~~$\Rightarrow $~~
A generator $g$ exists which produces $\mathbb{Z}_p^{*}$: $\mathbb{Z}_p^{*}=<g>$.\\

\noindent The following theorem from \cite[Pg. 69]{nt:Katzenbeisser2001}\index{Katzenbeisser 2001} characterizes all $(e-1)$-th roots of unity in $\mathbb{Z}_p^{*}$:
\begin{theorem}\label{nt-katzenbeisser-Anzahl-Einheitswurzeln}
$g^{\alpha}$ is exactly then $(e-1)$-th root of unity in $\mathbb{Z}_p^{*}$,
when $(e-1)\alpha = 0\bmod p-1.$ There are $gcd(p-1, e-1)$ of these.
\end{theorem}
\begin{proof} 
The first thesis results directly from the small theorem from Fermat:
\[g^{\alpha (e-1)}= 1 \bmod p  ~~\Rightarrow~~ \alpha (e-1) = 0\bmod p-1 \]
Let $\delta =gcd(p-1, e-1)$.  $\alpha (e-1) = 0\bmod p-1$ implies $\frac{\alpha (e-1)}{\delta}= 0 \bmod \frac{p-1}{\delta}$. \\
Since $\frac {e-1}{\delta}$ and $\frac{p-1}{\delta}$ are coprime (each was 
reduced by the gcd of their corresponding numerator),
$\alpha$ must be a multiple of $\frac{p-1}{\delta}$.

\[ \alpha \frac{p-1}{\delta} ~~ \text{with} ~~ \alpha = 1,...,\delta
\]\\
These $\delta$ different powers then correspond to the 
$(e-1)$-th roots of unity $g^{\alpha \frac{p-1}{\delta}} \bmod p$ in $\mathbb{Z}_p^{*}$.
\end{proof}

\noindent Analog for $q$: For $m^{e-1} = 1 \bmod q $ we then have $ gcd(q-1, e-1) $ many of $(e-1)$-th roots of unity.\\

\noindent The number of combinations of the $(e-1)$-th root of unity in $\mathbb{Z}_p^{*}$ and
$\mathbb{Z}_q^{*}$ gives the total quantity of RSA fixed points:
$ m^e = m \bmod n$ with $ m \in \{1,...,n-1\}$:\\
 $gcd(p-1, e-1) \cdot gcd(q-1, e-1) $\\

\noindent Adding $m=0$ to the above, results in the theorem \ref{nt-thm-Anzahl-RSA-Fixpunkte}:
\begin{theorem}\label{nt-thm-Anzahl-RSA-Fixpunkte}
If $ m \in \{0,...,n-1\}$, then the quantity of the RSA fixed points is:
          \[ (gcd(p-1, e-1)+1) \cdot (gcd(q-1, e-1)+1) \]
\end{theorem}
\end{proof}
\vspace{15pt}

% ----------------------------------------------------
\subsubsection{Lower bound for the quantity of RSA fixed points}
In the following section, we show that there is a lower bound for the quantity of RSA fixed points. This lower bound $6$ exists when the two different RSA prime numbers are the smallest possible values (2 and 3).\\

\noindent\textbf{Thesis 1: $p = 2, q = 3$}\\
The quantity of RSA fixed points for $p=2$ and $q=3$ is\\
$(\underbrace{gcd(p-1, e-1)}_{=1}+1) \cdot (\underbrace{gcd(q-1, e-1)}_{=2}+1)=2 \cdot 3=6$ \\

\noindent\textbf{Thesis 2: $p \neq q; p > 2, q > 2$}\\
The quantity of RSA fixed points for $p \neq q; p,q > 2$ is $\geq 9$.

\begin{proof}
Since $p$ and $q$ are prime, $(p-1)$ and $(q-1)$ for $ p,q > 2 $ are even.\\
The RSA algorithm requires to choose e so that $1 < e < \phi(n)=(p-1)(q-1)$ and \\
$gcd(e,(p-1)(q-1))=1$\\
Since $(p-1)$ and $(q-1)$ are even, e is odd $ \Rightarrow e-1$ is even.\\
Since $(p-1)$ and $(e-1)$ are even, then:\\
$gcd(p-1, e-1) \geq 2$ \\
$\Rightarrow (gcd(p-1, e-1)+1) \geq 3$ and $(gcd(q-1, e-1)+1) \geq 3$\\
$\Rightarrow (gcd(p-1, e-1)+1) \cdot (gcd(q-1, e-1)+1) \geq 9$
\end{proof}

\noindent Samples:\\
For $(e,n)=(17,6)$, all six possible messages \{0,1,2,3,4,5\} are
fixed points (for $n=6$, it is independent from the value of $e$).\\
For $(e,n)=(17,10)$, all 10 possible messages are fixed points.\\
For $(e,n)=(19,10)$, only 6 of the 10 possible messages are fixed points.

% ----------------------------------------------------
\vspace{15pt}
\subsubsection{Unfortunate choice of $e$ }
In this section, we show that with $e=1+lcm(p-1,q-1)$ each encryption results in a fixed
point (independently of the size of p,q, or n); and then we broaden this to all unfortunate
choices of $e$.\\

\noindent If $e=1$, then for all $ m$: $ c = m^e = m$. This is the trivial case.\\

\noindent\textbf{Thesis 1: $p,q > 2$}\\
If $e=1+lcm(p-1,q-1)$, then for all $ m \in \{1,...,n-1\}$: $ m^e = m \bmod n$.

\begin{proof}~\\
Given:\\
-~~~ $ ~~~ e\cdot d=1 \mod \phi(n)$ ~or~ $e\cdot d=1 \mod lcm(p-1,q-1) $\\
-~~~ $ ~~~ m^{x} \mod n = m^{x \mod \phi(n)} \mod n $\\

\noindent Encryption of messages: \\
 $~~~c=m^e \mod n$, where c is the ciphertext and m is the plaintext.

\noindent Decryption of messages:\\
 $~~~m'=c^d \mod n$, where d is the multiplicative inverse of e.\\

\noindent We will show: $c=m \mod n$ for the chosen e.

$~~~c = m^e \mod n$

$~~~c = m^{1+lcm(p-1,q-1)} \mod n$

$~~~c = m^1 \cdot m^{k \cdot (p-1) \cdot (q-1)} \mod n$

$~~~c = m^1 \cdot m^{[k \cdot \phi(n)] \mod \phi(n)} \mod n$

$~~~c = m^1 \cdot m^{0} = m \mod n $
\end{proof}

~\\
\noindent\textbf{Example: Fixed point property for all m:}\\
Given $n=p\cdot q= 13\cdot 37=481\\
\Rightarrow \phi(n)=(p-1)(q-1)=12\cdot 36=432$\\
$\Rightarrow e=lcm(p-1,q-1)+1=lcm(12,36)+1=36+1=37$.\\
With $m \in \{4,6,7,480\}$ we get in $m^{e} \mod n$ as:\\
$~~4^{37} \mod 481=~~4 $\\
$~~6^{37} \mod 481=~~6 $ \\
$~~7^{37} \mod 481=~~7 $ \\
$480^{37} \mod 481=480 $ \\   % be-xxxxxxxx Testen aller m-Werte auch mit 1 + 2*36 = 37+36= 73 !!!!!!!!!!!!!

~\\  
\noindent There is not just the one single $e$ (see above), where all
$ m \in \{1,...,n-1\}$ have the fixed point property $ m^e = m \bmod n$.\footnote{%
Sometimes these $e$, which make any message to a fixed point, are called a ``weak key'' $(e,n)$
of the RSA algorithm\index{Key!weak}.
This notation is different to the ``weak keys'' $k$ in DES\index{DES}, where \textbf{every}
message $m$ relates to itself if the \textbf{en}cryption is done twice.
To my knowledge, for larger $n$ the RSA procedure does not have weaks in this meaning:
$(m^e)^e = m$.}
\begin{theorem}\label{nt-thm-complete-fixed-point-property-values-of-e}
The complete fixed point property of all $m$ is valid for every
$e=j\cdot lcm(p-1,q-1)+1$, where $j=0,1,2,3,4, ... $ to $e \leq \phi(n)$.
\end{theorem}

~\\ % 
\noindent\textbf{Example: Further values for $e$ with fixed point properties:}\\
Given
$n=p\cdot q= 13\cdot 37=481$ with $lcm(p-1,q-1)=lcm(12,36)=36$.\\
Then, $e$ can have the following values: $e=j\cdot lcm(p-1,q-1)+1$ for $j=0,1,2,...,11$:\\
$\Rightarrow e \in \left\{ 1, 37,73,109,145,181,217, 253, 289, 325, 361,397\right\}$.\\

\noindent Starting $j=12$, the following is valid:
$ e=12\cdot lcm(12,36)+1=432+1=433 > 432=\phi(n)$.\\

\noindent Checking the four values above for $m$ with $e=217$, the results are:\\
$~~4^{217} \mod 481=~~4 $\\
$~~6^{217} \mod 481=~~6 $ \\
$~~7^{217} \mod 481=~~7 $ \\
$480^{217} \mod 481=480 $ \\

\begin{theorem}\label{nt-thm-complete-fixed-point-property-number-of-e}
The number of possible values for $e\text{ with } m^e = m \bmod n$
may be computed with the following:
\[\left[ \text{Quantity }e \right]=\left\lfloor \frac{\phi(n)}{lcm(p-1,q-1)+1}\right\rfloor+1 =\frac{\phi(n)}{lcm(p-1,q-1)}
\]
\end{theorem}

\noindent In our example, this results in $\frac{432}{lcm(12,36)}=12 $
different values for $e$, where $m^e = m \bmod n$ for all $m$ in $\mathbb{Z}_{481}$.\\
~\\ % Zeilenumbruch erzwingen

% ----------------------------------------------------
\subsubsection{An empirical estimate of the quantity of fixed points for growing moduli}
In this section, we make an empirical estimate of the quantity of fixed
points for growing moduli (and $e$ not weak).

\noindent For this, we randomly choose $p $ and $ q$ from the six following ranges
each characterized by its lower and upper bound:
$(2^2, 2^{10}), (2^{10}, 2^{20}), (2^{20}, 2^{40}), (2^{40}, 2^{80}),
 (2^{80}, 2^{160}), (2^{160}, 2^{320})$.\\
10 attempts were made for each range. For the exponent $e$, the standard value
$e=2^{16}+1$ was always chosen. The quantity of fixed points for all 60 attempts
was computed with the program \ref{nt_sagesample_Calculate_RSA-Fixpoints} ``Getfixpoints.sage''.

\noindent The following five sets contain the randomly chosen value pairs (p,q) of the first five ranges.

\begin{equation*}
  \begin{split}
From (2^2, 2^{10}): (p,q)~\in~ & \{ (127,947),(349,809),(47,461),(587,151),(19,23),\\ 
                              & (709,509),(653,11),(859,523),(823,811),(83,331)\} \\
\\
From (2^{10}, 2^{20}): (p,q)~\in~
            & \{ (447401,526283),(474223,973757),(100829,126757),\\
            &    (35803,116933), (577751,598783),(558121,607337),\\
            &    (950233,248167),(451103,73009),(235787,164429),\\
            &    (433267,287939)\}\\
  \end{split}
\end{equation*}

\begin{equation*}
  \begin{split}
From (2^{20},2^{40}): (p,q)~\in~
            & \{ (58569604997,321367332149),(286573447351,636576727223),\\
            &    (134703821971,134220414529),(161234614601,711682765579), \\
            &    (19367840881, 804790726361),(932891507377,521129503333),\\
            &    (337186437739,426034644493),(986529569219,604515928397),\\
            &    (276825557171,654134442649),(639276602353,1069979301731) \}\\
  \end{split}
\end{equation*}

\begin{equation*}
  \begin{split}
From (2^{40}, 2^{80}): (p,q)~\in~
            & \{ (667530919106151273090539,287940270633610590682889),\\
            &    (437090557112369481760661,590040807609821698387141),\\
            &    (1131921188937480863054851,813935599673320990215139)\\
            &    (874130181777177966406673,632270193935624953596331),\\
            &    (599303355925474677078809,717005631177936134003029),\\
            &    (752829320004631398659063,714134510643836818718761),\\
            &    (1046313315092743492917349,835721729660755006973833),\\
            &    (877161707568112212806617,42831503328261105793649),\\
            &    (575464819450637793425803, 5425832051159043433027),\\
            &    (321404337099945148592363,992663778486687980443879) \} \\
  \end{split}
\end{equation*}

\begin{equation*}
  \begin{split}
From (2^{80}, 2^{160}): (p,q)~\in~
   & \{ (838952969674957834783403492645269831354775774659,\\
   &     694309130163549038783972189350416942879771871411),\\
   &    (981985107290629501374187748859961786804311564643,\\
   &     178616495258601001174141825667078950281544628693),\\
   &    (614446632627716919862227545890890553330513965359,\\
   &     761232454374959264696945191327265643178491649141),\\
   &    (1421756952722008095585945863962560425554707936337,\\ 
   &     986781711714138924140285492105143175328486228197),\\
   &    (862346475785474165539441761205023498091366178341,\\ 
   &     438589995804600940885415547506719456975478582911),\\
   &    (1034081318899669345416602574034081247538053001533,\\
   &     1207032778571434704618111297072774884748706223447),\\
   &    (308083812465705343620096534684980088954958466893,\\ 
   &     350597371862294596793629011464584694618569736021),\\
   &    (830376326124356299120963861338027196931951857769,\\ 
   &     924874232653136669722297184352059466357375363191),\\
   &    (85600581120154590810189237569820706006659829231, \\
   &     297064381842806596646150718828138629443319259829),\\
   &    (1358984492013516052055790129324581847590275909129,\\
   &     609402294805414245544586792657989060761523960427) \} \\
  \end{split}
\end{equation*}

\begin{figure}[!htb]  %[htbp]
  \centering
   \includegraphics[width=0.92\textwidth]{figures/MAF.png}
  \caption{An empirical estimate of the quantity of fixed points for growing moduli}
  \label{fig:NumberFixpointsGrowingN}
  %\vskip +45pt
\end{figure}

Figure \ref{fig:NumberFixpointsGrowingN} shows that within the six ranges of size,
the average number of fixed points was not higher than 40.
% Unfortunately, we were unable to test the moduli for even higher ranges because 
% our Sage Program \ref{nt_sagesample_Calculate_RSA-Fixpoints}  "`Getfixpoints"'
% was too slow in ranges with $(p,q) > 2^{320}$.\\

\vspace{15pt}
% ----------------------------------------------------
\subsubsection{Example: Determining all fixed points for a specific public RSA key}

The exercise is to determine all fixed points for (n, e) = (866959, 17).\\

\noindent \textbf{Solution:} \\
We start by factoring $n$:  $ 866959 = 811 \cdot 1069$.\\

\noindent The quantity of RSA fixed points results from the theorem
\ref{nt-thm-Anzahl-RSA-Fixpunkte}:\\
$ (gcd(  p-1,  e-1)+1) \cdot (gcd(   q-1,  e-1)+1) =
  (gcd(811-1, 17-1)+1) \cdot (gcd(1069-1, 17-1)+1) = (2+1) \cdot (4+1) = 15 $\\

\noindent Sage program \ref{nt_sagesample_Calculate_RSA-Fixpoints} ``Getfixpoints''
returns the following 15 fixed points for $ (n,e) = (866959, 17)$:
\begin{table}[ht]
\begin{center}
{\tt 
\begin{tabular}{llllllll}
     0 &      1 &  23518 &  23519 &  47037 \\
188964 & 212482 & 236000 & 654477 & 843440 \\
843441 & 630959 & 677995 & 819922 & 866958 
\end{tabular} } % tt
\end{center}
\end{table}

\noindent \textbf{Example:} \\
Confirm that $ m = 843441$ is actually a fixed point:
$843441^{17} \mod 866959 = 843441 $\\

\begin{sagecode}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
import numpy

print "--- Search for fixpoints in Textbook-RSA given p, q, e ---";
fp=numpy.array([0])
fq=numpy.array([0])

#Edit e,p,q here
###EDIT BEGIN###
e=17;
p=811;
q=1069;
###EDIT END###

n=p*q;
print "Prime p: ",p;
print "Prime q: ",q;
print "Modul n: ",n;
print "Public exponent e: ", e;

r=Integers(p)
gen_f_p = r.multiplicative_generator(); print "\nGenerator of f_p: ",gen_f_p;
s=Integers(q)
gen_f_q = s.multiplicative_generator(); print "Generator of f_q: ",gen_f_q;

gcd_p = gcd(e-1,p-1)
gcd_q = gcd(e-1,q-1)
print "\ngcd(e-1,p-1): ", gcd_p;
print "gcd(e-1,q-1): ", gcd_q;

print "\nNumber of fixpoints: ",(gcd_p+1)*(gcd_q+1);
#Calculating fixpoints modulo F_p
#run i from 0 until gcd(e-1,p-1):
#g^( i*(p-1) / (gcd(e-1,p-1)) ) mod p

print "\nFixpoints modulo p";
print "0 (trivial fixpoint added manually)";
i=0;
for i in range(gcd_p):
                fix_p = power_mod(gen_f_p,Integer(i*(p-1)/gcd_p),p); print fix_p;
                fp = numpy.append(fp,fix_p)

print "\nFixpoints modulo q";
print "0 (trivial fixpoint added manually)";
j=0;
for j in range(gcd_q):
                fix_q = power_mod(gen_f_q,Integer(j*(q-1)/gcd_q),q); print fix_q;
                fq = numpy.append(fq,fix_q);

print "\nFixpoints for the public RSA key (n,e) = (", n, ",", e, ")"
for r in fp:
       for s in fq:
               print crt(Integer(r),Integer(s),Integer(p),Integer(q))

print "\nRemark: You can verify each fixpoint with power_mod(m,e,n).";
\end{Verbatim}
\caption{Determining all Fixed Points for a Specific Public RSA Key}
\label{nt_sagesample_Calculate_RSA-Fixpoints}
\end{sagecode}
%#print "        Here done for the last found fixpoint:";
%#m = crt(Integer(r),Integer(s),Integer(p),Integer(q))
%#print "        m = ", m, ",", "power_mod = ", power_mod(m,e,n)
%#if (m != power_mod(m,e,n)):
%#                print "Verification failed !!!";

\vspace{60pt}
\noindent \textbf{Meaning of the Variables in the Sage Code
                  \ref{nt_sagesample_Calculate_RSA-Fixpoints}:}
\begin{Verbatim}%
[fontsize=\footnotesize,fontshape=tt]
- gen_f_p = r.multiplicative_generator()
  r is a residue class ring modulo p and multiplicative_generator() returns
  a generator element that was created by the ring modulo p.
- power_mod(gen_f_p,Integer(i*(p-1)/gcd_p),p)
  The power_mod function raises a number m to the power of e and returns the results modulo n.
  E.g.: power_mod(m, e, n) := m^e modulo n
- numpy.append(fp,power_mod(gen_f_p,Integer(i*(p-1)/gcd_p),p))
  The append function extends an array (fp) by an additional element.
- crt(Integer(r),Integer(s),Integer(p),Integer(q))
  CRT is the acronym for Chinese Remainder Theorem. crt(r, s, p, q) solves
  the congruences x = r mod p and x = s mod q with the help of the Chinese Remainder Theorem.
\end{Verbatim}

%%% xxx111222xxx-end

@@

\begin{center}
\begin{tabular}{|l|l|l|}\hline
 & Short description ~~ & Page \\ \hline

Definition~\ref{def-zth-prime} & prime numbers &  \pageref{def-zth-prime} \\
Definition~\ref{def-zth-composite} & composite numbers & \pageref{def-zth-composite}  \\ \hline

Theorem~\ref{thm-zth-cnum} & factors of composite numbers~~~~~~~ & \pageref{thm-zth-cnum}\\
Theorem~\ref{thm-zth-mthm} &  1st fundamental theorem of number theory &  \pageref{thm-zth-mthm} \\  \hline

Definition~\ref{def-zth-divisibility} & divisibility & \pageref{def-zth-divisibility} \\
Definition~\ref{def-zth-remainder} & remainder class $r$ modulo $m$ & \pageref{def-zth-remainder} \\
Definition~\ref{def-zth-congruence} & congruent & \pageref{def-zth-congruence} \\ \hline

Theorem~\ref{thm-zth-div} & congruence with difference  & \pageref{thm-zth-div} \\
Theorem~\ref{thm-zth-multinv} & multiplicative inverse (existence) & \pageref{thm-zth-multinv}  \\
Theorem~\ref{thm-zth-exhperm} & exhaustive permutation & \pageref{thm-zth-exhperm} \\
Theorem~\ref{thm-zth-pot} & power mod $m$ & \pageref{thm-zth-pot} \\ \hline

Definition~\ref{def-zth-zn} & $\mathbb{Z}_n$  & \pageref{def-zth-zn}\\
Definition~\ref{def-zth-znmult} &   $\mathbb{Z}_n^*$ & \pageref{def-zth-znmult} \\ \hline

Theorem~\ref{thm-zth-znmult} & multiplicative inverse in $\mathbb{Z}_n^*$& \pageref{thm-zth-znmult} \\ \hline

Definition~\ref{def-zth-phiofn} & Euler function $J(n)$ & \pageref{def-zth-phiofn} \\
Theorem~\ref{thm-zth-phiprime} & $J(p)$ &  \pageref{thm-zth-phiprime}\\
Theorem~\ref{thm-zth-phipq} & $J(p*q)$ &  \pageref{thm-zth-phipq}\\
Theorem~\ref{thm-zth-phimultprime} & $J(p_1 * \cdots *p_k)$ & \pageref{thm-zth-phimultprime} \\
Theorem~\ref{thm-zth-phinum} & $J(p_1^{e_1} * \cdots *p_k^{e_k})$ & \pageref{thm-zth-phinum} \\
Theorem~\ref{thm-zth-fermat1} & little Fermat  &  \pageref{thm-zth-fermat1}\\
Theorem~\ref{thm-zth-fermateuler} & Euler-Fermat theorem & \pageref{thm-zth-fermateuler} \\ \hline

Definition~\ref{def-zth-ordn} & multiplicative order $ {\rm ord}_{m} (a)$ & \pageref{def-zth-ordn} \\
Definition~\ref{def-zth-primitiveroot} & primitive root of $m$ &  \pageref{def-zth-primitiveroot}\\
Theorem~\ref{thm-zth-ordp} & exhausting of all possible values & \pageref{thm-zth-ordp} \\ \hline

Theorem~\ref{nt-thm-Anzahl-RSA-Fixpunkte} & number of RSA fixed points & \pageref{nt-thm-Anzahl-RSA-Fixpunkte} \\ \hline

\end{tabular}
\end{center}
\vskip +6 pt

@@

\bibitem[Agrawal2002]{nt:Agrawal2002}  \index{Agrawal 2002} 
    M. Agrawal, N. Kayal, N. Saxena, \\
    {\em PRIMES in P}, August 2002, revised paper: \\
       \url{http://www.cse.iitk.ac.in/~manindra/algebra/primality_v6.pdf}\\
    See also the website "The AKS "PRIMES in P" Algorithm Resource":\\
       \url{http://fatphil.org/maths/AKS/}.

@@

\bibitem[Bourseau2002]{nt:Bourseau2002} \index{Bourseau 2002} \index{Fox 2002}
    F. Bourseau, D. Fox, C. Thiel, \\
    {\em Vorz\"uge und Grenzen des RSA-Verfahrens},\\
    In: Datenschutz und Datensicherheit (DuD) 26/2002, pp~84-89 (see www.dud.de),\\
    \url{http://www.secorvo.de/publikationen/rsa-grenzen-fox-2002.pdf}.

@@

  \item {Henrik Koy for making many very useful suggestions, 
         for the very constructive proof-reading of the first version of
	 this article, and for helping with TeX. \\
	 Mr.\ Koy designed and developed in his leisure time the 
	 functions and the complex dialog box of the RSA cryptosystem within CrypTool 1, 
	 which enables you to execute the RSA samples of this article.}

@@

  \item {Minh Van Nguyen from Australia for his always fast, professional and exhaustive
         help with the first Sage\index{Sage} code samples in this chapter.}
